---
title: "knobi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{knobi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(knobi)
```

```{r,echo=FALSE}
library(icesSAG)
```

&nbsp;


For the correct understanding of the Known biomass production models (KBPMs) models we start reviewing the 
surplus production models (SPMs) framework. Then, we focus in the KBPMs formulation to provide the required knowledge for a correct package use.


&nbsp;

Traditional SPMs are one of the most used assessment models for data-limited  marine populations
Their general structure relates directly to Russell’s formulation of the stock dynamics, and they have the following general structure,

$$B_{t+1}=B_t + f(B_t)-C_t$$
where *B<sub>t</sub>* is the stock biomass, *C<sub>t</sub>* is the biomass caught and *f(B<sub>t</sub>)* is the biomass production function. The subscript *t* denotes the time (years).

There are many formulations of the production of biomass function *f(B<sub>t</sub>)*, among which the general Pella-Tomlinson is widely used:

$$f(B_t) = \frac{r}{p}{B_{t}} \left(1-\left( \frac{B_{t}}{K}\right) ^{p}\right)$$ (1)

where *r* is the intrinsic population grown rate, *K* is the virgin biomass and *p* is the asymmetry parameter, used so that the production curve is not always symmetrical and consequently its maximum production is not always produced in *K/2*.

These SPMs relates historical series of catch to historical fishing effort or indexes of relative biomass such as CPUE (catch-per-unit-effort). For this purpose, such SPMs require the estimation of a catchability coefficient (*q*) to relate the biomass index to an (inferred) true biomass. 

$$\hat{I}_t=C_t/E_t=qB_t$$
where *I<sub>t* is an index of relative biomass for year *t*, notation *ˆ* denotes an estimated value and *q* is the catchability coefficient, which scales the modelled stock biomass to match the trends in catch rates.

An alternative line of research based on surplus production models named known-biomass production models (KBPM) was developed (MacCall, 2002) based on the idea that the annual surplus production in an unfished stock is equal to *B<sub>t+1</sub>-B<sub>t*, and that, for a fished stock, the calculation of surplus production depends on catch. 

$$SP_t=\overline{B}_{t+1}-\overline{B}_t+C_t$$ (2)

where *SP<sub>t</sub>* is the surplus production, *<SPAN STYLE="text-decoration:overline">B</SPAN><sub>t</sub>* is the average biomass or SSB (mean of two consecutive years), *<SPAN STYLE="text-decoration:overline">B</SPAN><sub>t</sub>=(B<sub>t</sub>+B<sub>t+1</sub>)/2*, and *C<sub>t</sub>* represent the catch. The subscript *t* denotes the time (years). 

In contrast to the traditional SPMs, KBPMs use as input data a biomass time series produced by other stock assessment model instead of biomass index and thereby avoid the imprecision associated with estimating the catchability coefficient. 

Then, surplus production is calculated from the known average biomass (of two consecutive years) and the observed catch using equation (2). Then, for the KBPMs fit, equation (1) is used:

$$SP_{t}= \frac{r}{p}\overline{B}_{t}\left(1-\left( \frac{\overline{B}_{t}}{K}\right) ^{p}\right)$$ (3)

Schaefer model corresponds to *p=1* (then the production curve is symmetric and *SP<sub>max</sub>* is equal to *K/2*).

&nbsp;


## `knobi_fit` ##

`knobi_fit(data, control, plot_out = F, plot_filename = NULL, plot_dir = NULL)`

&nbsp;

#### Creating `data` argument ####

In this section the use of the knobi_fit is illustrated using the northern hake (*Merluccius merluccius* ) as a case of study.

The first step is to create the data input object using the available data for our case study. Mandatory data for the known-biomass production model (KBPM) are the biomass or the spawning stock biomass (SSB) time series, the catch time series and the corresponding years.  

In this case, we use the available data for European hake in the International Council for the Exploration of the Sea (ICES). The northern hake stock covers the subareas 4, 6, and 7, and divisions 3.a, 8.a–b, and 8.d (Greater North Sea, Celtic Seas, and the northern Bay of Biscay). The data time period goes from 1978 to 2020.  

The data is downloaded using `icesSAG` package and saved `hake_n` object. Then the data list for `knobi_fit` is created.

```{r}
hake_n <- icesSAG::getSAG(stock = "hke.27.3a46-8abd", year = 2021)
hake_n <- hake_n[-nrow(hake_n),]      # we remove last line because there is 
                                      # no information about catch in the last year

```



```{r}
head(hake_n,n=3)
tail(hake_n,n=3)

data<-list()
data$Biomass=hake_n$biomass           # Biomass or
data$Spawning_Biomass=hake_n$SSB      # Spawning stock biomass
data$Catch=hake_n$catches             # Catch
data$F_input=hake_n$F                 # Fishing mortality (optional)
data$Recruitment=hake_n$recruitment   # Recruitment (optional)
data$RP=list(F_MSY=0.26)              # Reference points (optional)
                                      # in this case, Fmsy is provided by ICES        
data$classF_input="average"           # Fishing mortality class (optional)
data$years=hake_n$Year                # Years
```

&nbsp;

#### Creating `control` object ####

&nbsp;

Control list contains a set of settings for the KBPM fit.

Below, we provide values for the mandatory ones.

&nbsp;

```{r}
control=list()
control$pella="TRUE"            # "TRUE" for Pella-Tomlinson model or "FALSE" for Schaefer (by default)
control$method="SSB"                # "SSB" or "Biomass" according to the data
                                    # (required if both time series are provided)
```

&nbsp;

After preparing both lists, `data` and `control`, we can apply the `knobi_fit` over them.

```{r,eval=FALSE}
hake_n_results<-knobi_fit(data,control,
                          plot_out=T,                 # This argument allows us to create, in addition 
                                                      # to the internal plots of R, an external folder
                                                      # with the corresponding files. FALSE by default.
                          plot_filename="hake_n",     # Optional name of the folder if plot_out=T. 
                                                      # "knobi_results" by default
                          plot_dir="Z:/knobi/HAKE/")  # Optional folder directory (current one by default) 
```


Note that if the length of the input catch time series does not match with the biomass or SSB length, a warning is returned indicating that the series of catch is reduced so that the fit can be done.

Furthermore, the plots are shown in the plot window and also saved (if `plot_out=TRUE`) on the provided directory or in the current directory (if the directory is not provided). Plots of the following input quantities are shown: time series of fishing mortality; SSB or biomass according to the data entered or the chosen method; the surplus production (SP) calculated from the equation explained above; and catch time series. Next plots show us the catch versus fishing mortality and fishing mortality over SSB if the information about fishing mortality is provided, and also the relation between catch over SSB. After that, the fitted surplus production curve is plotted with the SSB and SP observations and also with the catch and SP observations. The last four plots show a comparison between the KBPM and the ICES (in our case, provided by Stock Synthesis model) estimates of fishing  mortality and SSB, in absolute and relative (*F/F<sub>MSY* and *SSB/SSB<sub>MSY*, respectively) terms. In those plots that show absolute quantities, the horizontal lines represent the ICES (if they are provided) and the KBPM estimates of *F<sub>MSY* and *SSB<sub>MSY*. Fishing mortality is calculated as *F<sub>t</sub>=C<sub>t</sub>/<SPAN STYLE="text-decoration:overline">B</SPAN><sub>t*.

```{r,echo=FALSE, fig.width=8, fig.height=6}
hake_n_results<-knobi_fit(data,control,plot_out=F)
```

&nbsp;

#### Quantitative results ####

The output is a list containing the input data and control updated with the fitting results and the plot settings information, respectively. The data is updated including also the annual average biomass (mean of two consecutive years), in `\$data\$Average_Biomass`, the surplus production, in `$data\$SP`, and the *F* estimates derive from KBPM, in `\$data\$F_output`. The fit results of the KBPM are in `\$fit`, and contains:

* `Parameter_estimates`: Estimates of the model parameters.
* `data`: Reports the data used for the model.
* `RP`: Estimates of the biological reference points. 
  + `K`: KBPM estimate of virgin biomass (*K*).
  + `B_MSY`: KBPM estimate of biomass at maximum sustainable yield.
  + `F_MSY`: KBPM estimate of fishing mortality at maximum sustainable yield.
  + `MSY`: KBPM estimate of maximum sustainable yield.
  + `MSYoverK`: ratio of MSY and K.
* `optimr`: Information about the `optimr` function results
  + `value`: The value of the function corresponding to the parameter estimation.
  + `convergence`: An integer code. ‘0’ indicates successful completion in the optimization.
  + `message`: A character string giving any additional information returned by the optimizer, or NULL.
* `error`: Data frame with measures of estimates accuracy (error measures comparing observed an estimated values) and model performance: 
  + `SER`: Standard error of the regression.
  + `R-squared`: Coefficient of determination.
  + `adj-R-squared`: Adjusted coefficient of determination.
  + `AIC`: Akaike information criterion.
  + `RMSE`: Root mean squared error.
  + `MAPE`: Mean absolute percentage error.}



```{r}
hake_n_results$fit
```

&nbsp;

Comparisons of ICES and KBPM estimates can be done. For example, below we compare the *F<sub>MSY* estimates.

```{r,echo=FALSE}
refpts<-cbind(hake_n_results$data$RP$F_MSY,hake_n_results$fit$RP$F_MSY)
colnames(refpts)<-c("SS_input_RP","knobi_output_RP")
rownames(refpts)<-c("F_msy")
options(scipen=999)
round(refpts,3)
```



&nbsp;

## `knobi_retro` ##

Once the KBPM fit is carried out using `knobi_fit`, its robustness to the systematically deletion of last year of data is tested using the `knobi_retro`.


&nbsp;

`knobi_retro` input is the object returned by `knobi_fit` and the number of retrospective patterns. Note that `nR` specifies the number of fits to carry out. The first model considers the data deleting the last year and fits the surplus production curve, the next model deletes the two last years of the original data set and fits the SP curve, and then continues in this way until the last model is reached in which the last `nR` years in the original data are deleted to then fit the curve.

```{r,eval=FALSE}
hake_n_retros<-knobi_retro(hake_n_results,             # previous results
                                   nR=5,               # number of patterns
                                   plot_out=T,
                                   plot_filename="hake_n",
                                   plot_dir="Z:/knobi/HAKE/")
```

```{r,echo=FALSE, fig.width=8, fig.height=6}
hake_n_retros<-knobi_retro(hake_n_results,nR=5,plot_out=F)
```

&nbsp;

#### Quantitative results ####

A list containing the retrospective analysis results is returned. It includes the fits and the list of reference points.

```{r}
hake_n_retros
```

&nbsp;

There is also the possibility of choosing the time series of the retrospective analysis, through the arguments:
* `yR`: Vector containing the last years considered in the analysis.
* `yR0`: Optional vector with the first years corresponding to `yR`.

```{r, fig.width=8, fig.height=6}
knobi_retro(hake_n_results,yR=c(2005,2010,2015),plot_out=F)
```

&nbsp;

```{r, fig.width=8, fig.height=6}
knobi_retro(hake_n_results,
                           yR= c(2005,2010,2015),
                           yR0=c(1990,1995,1995),
                           plot_out=F)
```
&nbsp;
&nbsp;

## `knobi_env` ##

After carrying out the KBPM fit using `knobi_fit`, `knobi_env` function allows us to analyse and model the relationships between surplus production and environmental covariables to test whether productivity changes in response to environmental fluctuations. Environmental effects are included as additive and multiplicative effects in the general KBPM formulation (3).

* Additive model:

$$SP_{t}= \frac{r}{p}\overline{B}_{t}\left(1-\left( \frac{\overline{B}_{t}}{K}\right) ^{p}\right) + cX_t\overline{B}_{t}$$ (4)
being *c* the parameter that represent the effect of the environmental variable *X<sub>t</sub>* (*t* index represents years).

* Multiplicative model:

$$SP_{t}=  \frac{r}{p}\overline{B}_{t}\left(1-\left( \frac{\overline{B}_{t}}{K}\right) ^{p}\right)exp^{cX_t}$$ (5)


&nbsp;

`knobi_env` input is the object returned by `knobi_fit` and the object `environmental`, a list containing the environmental information required for the fit. In this example, we provide the mandatory information, which is:

* `data`: data frame containing the values of each of the environmental variable(s) in one column. Each row represents a year.
* `years`: time series of years corresponding to the environmental variable(s).
* `nlag`: this argument is used to test all the lags smaller or equal to `nlag` (numerical vector) through cor.test function. The lag corresponding to the highest pearson correlation among the surplus production and the lag environmental covariable values is considered in the environmental model. This means that the *SP<sub>t</sub>* is related to *X<sub>t-lag</sub>* being *X* the environmental variable. The length of this argument must be equal to the number of environmental variables included.

```{r}
Env=data.frame(years=seq(1973,2020),AMO=c(-0.236,-0.441,-0.32,-0.385,-0.21,-0.201,-0.132,
  -0.041,-0.098,-0.235,-0.093,-0.23,-0.29,-0.297,0.044,-0.028,-0.106,-0.061,-0.155,-0.242,
  -0.234,-0.2,0.112,-0.082,0.028,0.349,0.094,0.004,0.095,0.041,0.207,0.182,0.268,0.242,
  0.123,0.114,0.015,0.325,0.078,0.189,0.142,0.077,0.09,0.318,0.291,0.045,0.15,0.279),
  TMax_Vigo=c(16.5,16.7,17.1,16.3,16.4,16.7,17.3,17.3,17.4,18,17.5,17.5,17.6,18,17,18.4,
  17.8,19.6,19.1,18,17.9,17.7,17.7,19.4,18.2,19.7,19.2,18.6,18,18.3,18.5,18.8,18.7,18.7,
  18.5,17.9,17.4,19.2,19,19.7,18,19.1,19.4,20,19.5,20.2,18.8,18.6))

# The environmental data series must start in the first year of the KBPM fit data
# minus the provided nlag or lag
years=hake_n_results$data$years # See knobi_fit example to obtain the knobi_results object
ind=which(Env[,1]==years[1])
ind1=which(Env[,1]==years[length(years)])
nlag=5
Env=Env[(ind-nlag):ind1,]

# Now we create the environmental list
environmental=list()
environmental$data=data.frame(AMO=Env$AMO,Tmax=Env$TMax_Vigo)
environmental$years=Env$years
environmental$nlag=c(nlag,nlag)
```

```{r,eval=FALSE}
hake_n_environmental<-knobi_env(hake_n_results,        # previous results
                                environmental,         # the environmental information
                                plot_out=T,
                                plot_filename="hake_n",
                                plot_dir="Z:/knobi/HAKE/"))
```


```{r,echo=FALSE, fig.width=8, fig.height=6}
hake_n_environmental<-knobi_env(hake_n_results,environmental)
```

Result plots are shown in the plot window and also saved (if `plot_out="TRUE"`) on the provided directory or in the same directory as `knobi_fit`.
 
The first plot reports the correlation analysis between the environmental variable(s) and the KBPM data. The second one reports the fitted values of the base model (no environmental information) and of the environmental ones.
 
3D plots reporting the surplus production curve conditioned to a grid of environmental values are also reported.

&nbsp;

#### Quantitative results ####

The output is a list containing the environmental analysis:

* `selected_lag`: Estimated lag corresponding to the one reporting the highest correlation between the environmental variable and the surplus production. Derived if lag is not fixed.
* `lag_cor`: Correlation between the environmental variable(s) value and the surplus production (after lagging the environmental one if corresponds).
* `selected_var`: Environmental variable used in the fit, chosen by the user or the one derived from the highest pearson correlation procedure.
* `model_env_Multiplicative`: Estimates of the multiplicative model parameters.
* `model_env_Additive`: Estimates of the additive model parameters.
* `ref_pts`: Reference points (RPs) estimations for each model in the absence of environmental effect.

From equations (4) and (5), estimations of the most used reference points are shown, but is has to be considered that RPs are dependent on the environmental covariable value, and then their values varies according to the environmental covariable values. Hence, for simplicity, the output reference points consider that the environmental covariable is 0. 

In the case of the multiplicative model, the RPs that depends on the environmental covariable are *F<sub>msy</sub>* and *MSY*. If the user wants to know their values at a specific environmental covariable value, please multiply this RPs by *cX<sub>t*. 

In the case of the additive model, all RPs depend on the environmental covariable values. If the user wants to know their values at a specific environmental covariable value, consider the following equations:

$$B_{msy}(X_t)=K\left(\frac{p c X_t+r}{r(p+1)}\right)^{1/p}$$
$$F_{msy}(X_t)=F_{msy}-\frac{cX_t}{p+1}+cX_t$$
$$MSY(X_t)=B_{msy}(X_t)*F_{msy}(X_t)$$
$$K(X_t)=K+cX_t$$
*F<sub>msy</sub>*, *B<sub>msy</sub>* and *K* are the RP estimates for *X<sub>t</sub>=0*. 
Note that *RP(X<sub>t</sub>)* represents the dependence of each RP in relation to the environmental variable.

* `scaled_environmental_var`: Standardized variable used in the fit, with the `scale` and `center` attributes.
* `plots3D`: List with the 3D plots objects.
* `error`: Array of  measures of estimates accuracy (error measures comparing observed an estimated values) for each model (base KBPM, additive model and multiplicative model): Standard error of the regression (SER), coefficient of determination (R-squared), adjusted coefficient of determination (adj-R-squared), Akaike information criterion (AIC), root-mean-squared error (RMSE), mean absolute percentage error (MAPE) and the value of the F statistic corresponding to the comparison of each environmental model respect to the base model (F-value) and its corresponding p-value (Pr(>F)).

```{r, fig.show='hide'}
hake_n_environmental
```

&nbsp;
&nbsp;

#### More options ####

When more than one environmental variable is introduced in `data`, the fit is made with the most correlated environmental variable considering the selected lag, but there is the possibility of choosing one of the introduced variables through the argument `selected_var` inside the `environmental` list. 

```{r, eval=FALSE}
environmental$selected_var="Tmax"
knobi_env(hake_n_results,environmental)
```

&nbsp;

There is the possibility of fixing lag value(s) in the relation among the surplus production and the environmental variable, using `lag` instead `nlag` inside `environmental`.

It is also possible to fit the environmental model considering more environmental variables, up to a total of 5 covariates, using `multicovar = TRUE`. This means that *cX<sub>t</sub>* is replaced by *sum<sub>{i in 1...n}</sub>(c<sub>i</sub>X<sub>t,i</sub>)* in equations (4) and (5), where index *n* represents each environmental variable.

```{r, eval=FALSE}
environmental=list()
environmental$data=data.frame(AMO=Env$AMO,Tmax=Env$TMax_Vigo)
environmental$years=Env$years
environmental$lag=c(2,3)
environmental$multicovar=TRUE
hake_n_multi<-knobi_env(hake_n_results,environmental)
```

In this case, 3D plots are not reported, and the output list slightly differs from the previous one.

* `fixed_lag` instead `selected_lag`, which is the input value of 'lag' argument.
* `environmental_variables` instead `scaled_environmental_var`, the standardized covariables used in the fit with the `scale` and `center` attributes.
* `plots3D` and `selected_var` are not reported.

For the RP estimation when *X<sub>t* is not equal to 0, you have to replace the term *cX<sub>t* in the previous *RP(X<sub>t</sub>)* equations by *sum<sub>i in 1...n</sub>(c<sub>i</sub>X<sub>t,i</sub>)*, where *n* is the number of covariates.




&nbsp;
&nbsp;

## `knobi_proj` ##

`knobi_proj` function allows us to projects the stock biomass  (or stock spawning biomass) time series and then the surplus production  based on the selected catch or fishing mortality values for the projected years after carrying out the KBPM fit using `knobi_fit` or the environmental KBPM fit using `knobi_env`.

&nbsp;

`knobi_proj` input is the object returned by `knobi_fit` and the following information:

* `env_results`: The output object of `knobi_env` function. This argument is optional. It has to be provided if you also want to consider the environmental parameter estimates for the projections.
* `n_y`: The number of years for projections. If not provided, it has to be provided `end_y` (the next argument).
* `end_y`: The end year of the projections. If not provided, it has to be provided `n_y` (the argument above). If both are provided, the function uses n_y.
* `Ct`: Optional vector, data frame or matrix containing the values of the catch for the projected. Different catch scenarios are allowed. The values for each should be provided in each of the columns. The length vector or row number should be equal to the number of years. Projections can be based on selected catch or fishing mortality values, then only one of the arguments, `Ct` or `f`, must be introduced.
* `f`: Optional vector, data frame or matrix containing the values of the fishing mortality for the projected. Different catch scenarios are allowed. The values for each should be provided in each of the columns. The length vector or row number should be equal to the number of years. Projections can be based on selected catch or fishing mortality values, then only one of the arguments, Ct or f, must be introduced.
* `env`: This argument has to be provided if `env_results` is provided too. If the `multicovar` argument of `knobi_env` is FALSE, `env` has to be a vector, data frame or matrix containing the values of the environmental covariates (unstandardized) for the projection years (rows) and the different catch or fishing mortality settings (columns).  On the other hand, if the `multicovar` argument of `knobi_env` is TRUE, the current argument must be a list, and each argument must be a data frame corresponding to each catch or fishing mortality setting containing the values of the environmental covariates for that scenario.


```{r}

# First step is to create the data frame containing the selected catch for the projected years.
# In this illustration, catch values for all the projected years are constant and equal to the 
# catch value in the last year multiplied by 1, 1.2 and 0.8.

catch=rep(hake_n_results$data$Catch[length(hake_n_results$data$Catch)],5)

Ct=data.frame(catch=catch,
              catch08=0.8*catch,
              catch12=1.2*catch)


# Then, create the data frame containing the environmental covariable values
# for the projected years.

env=data.frame(AMO1=c(0.2,0.3,0.4,0.5,0.6),
               AMO2=c(0.05,0.15,0.25,0.35,0.45),
               AMO3=c(-0.1,-0.2,-0.3,-0.4,-0.5))
```



Based on the previous objects we can apply the projection function.

```{r}
knobi_proj(hake_n_results, hake_n_environmental, Ct=Ct, n_y=5, env=env)
```

The plots results are displayed in the plot window and are also saved (if `plot_out="TRUE"`) in the  provided directory or in the same directory as `knobi_fit`.
  
If environmental information is not provided, four plots are presented in a panel reporting the biomass, surplus production, catch and fishing mortality projections for each catch or fishing mortality scenario.  

If environmental information is provided, the plots are presented in a panel for each catch or fishing mortality scenario reporting biomass, surplus production, catch and fishing mortality projections in each of the environmental scenarios.
 
&nbsp;

#### Quantitative results ####

The output is a list containing the projection estimates:

* `base_model`: Three-dimensional array containing projection information for each catch or fishing mortality scenario that is projected using the KBPM base model. Rows correspond to years, columns to the derived quantities (biomass or SSB, surplus production, F and catch) and third dimension to the projection scenarios.
* `additive_model`: Four-dimensional array containing projections information for each catch or fishing mortality scenario and for each environmental scenario considering the KBPM additive model. The rows correspond to the years, the columns to the derived quantities (biomass or SSB, surplus production, F and catch), the third dimension to the environmental projection scenarios and the fourth one to the catch scenarios. It is only returned if the arguments env_results and env are provided.
* `multiplicative_model`: Four-dimensional array containing projections information for each catch or fishing mortality scenario and for each environmental scenario considering the KBPM multiplicative model. The rows correspond to the years, the columns to the derived quantities (biomass or SSB, surplus production, F and catch), the third dimension to the environmental projection scenarios and the fourth one to the catch scenarios. It is only returned if the arguments env_results and env are provided.
* `biomass`: Data frame containing historical biomass and their projections for all scenarios.
* `catch`: Data frame containing historical catch and their projection values for all scenarios.
* `f`: Data frame containing the historical fishing mortality and their projection values for all scenarios.
* `SP`: Data frame containing the historical surplus production and their projection values for all scenarios.

&nbsp;
&nbsp;

#### More options ####

Another example without environmental information for the projections is given below. The number of projection years is also extended.

```{r, eval=FALSE}
Ct=rbind(Ct,Ct[1:3,])
knobi_proj(hake_n_results, Ct=Ct, end_y=2027)
```

Alternatively, projections can be based on fishing mortality.
The scenarios presented below have been created from the estimated *F<sub>msy</sub>* of `knobi_fit` analysis.

```{r, eval=FALSE}
fmsy=hake_n_results$fit$RP$F_MSY
ff=rep(fmsy,5)
f=data.frame(f=ff,f12=ff*1.2,f08=ff*0.8)

knobi_proj(hake_n_results, f=f, n_y=5, env_results=hake_n_results, env=env)
```

The example without environmental information for the projections is given below. The number of projection years is also extended.

```{r, eval=FALSE}
f=rbind(f,f[1:3,])
knobi_proj(hake_n_results, f=f, end_y=2027)
```

In case of `multicovar=T` in `knobi_env`, a list is required which each item is a data frame for each environmental scenario.

```{r, eval=FALSE}
env=list(climate_1=data.frame(AMO=c(0.2,0.2,0.3,0.3,0.4),
                              Tmax_Vigo=c(19,19,20,20,21)),
         climate_2=data.frame(AMO=c(0.2,0.3,0.4,0.5,0.6),
                              Tmax_Vigo=c(19,20,21,22,23)))

knobi_proj(hake_n_results, hake_n_multi, Ct=Ct[1:5,], n_y=5, env=env)
```

## References ##

MacCall, Alec. (2002). Use of Known-Biomass Production Models to Determine Productivity of West Coast Groundfish Stocks. North American Journal of Fisheries Management. 22. 272-279. DOI: 10.1577/1548-8675%282002%29022%3C0272%3AUOKBPM%3E2.0.CO%3B2.

